---
title: "Take-Home Exercise 1.0: Creating Data Visualisation Beyond Default"
author: "Sheryl Ann Tan Yi-Shi"
date: "2025-02-15"
date-modified: "last-modified"
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1.0 Introduction

## 1.1 Background

Cardiovascular diseases, including heart attacks, are a leading cause of mortality worldwide. In recent years, Japan has experienced a steady increase in heart disease-related mortality, largely driven by its aging population, making it a growing public health concern (Ohira et al., 2023).

## 1.2 The Task

This exercise involves assuming the role of a graphical editor at an international media company that regularly publishes content on digital platforms. The company plans to release articles focused on one of the following themes:

-   Heart Attack in Japan

-   Ship Performance in the Gulf of Guinea

For the purposes of this take-home exercise, the theme, ***Heart Attack in Japan***, was selected for the preparation of data visualizations for the article.

## 1.3 About the Dataset

The dataset used in this exercise is sourced from Kaggle and can be accessed via this [link](https://www.kaggle.com/datasets/ashaychoudhary/heart-attack-in-japan-youth-vs-adult).

This dataset provides the opportunity to perform a comprehensive analysis of heart attack incidents in Japan as it includes a range of features such as:

-   **Age and Demographics** including gender and region.

-   **Medical Indicators** including cholesterol level, heart rate, systolic blood pressure, and diastolic blood pressure.

-   **Physical Metrics** including Body Mass Index (BMI) measurements.

-   **Behavioural Factors** including physical activity levels and levels of alcohol consumption.

# 2.0 Data Preparation

## 2.1 Loading R Packages

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

## 2.2 Importing the Data

```{r}
exam_data <- read_csv("data/japan_heart_attack_dataset.csv")
```

# 3.0 Data Visualisation

# 7.0 References

Ohira, T., Eguchi, E., Hayashi, F., Kinuta, M., & Imano, H. (2023). Epidemiology of cardiovascular disease in Japan: An overview study. *Journal of Cardiology, 81*(5), 379-387. https://doi.org/10.1016/j.jjcc.2023.07.007
