---
title: "Hands-on Exercise 4.4: Funnel Plots for Fair Comparisons"
author: "Sheryl Ann Tan Yi-Shi"
date: "2025-02-01"
date-modified: "last-modified"
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1.0 Getting Started

Funnel plot is a specially designed data visualisation for conducting unbiased comparison between outlets, stores or business entities.

## 1.1 Installing and Loading R Packages

The following R packages would be used in this exercise:

| R Package     | Description                              |
|---------------|------------------------------------------|
| `readr`       |                                          |
| `FunnelPlotR` | For creating funnel plot                 |
| `ggplot2`     | For creating the funnel plot manually.   |
| `knitr`       | For building the static html table.      |
| `plotly`      | For creating an interactive funnel plot. |

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

## 1.2 Importing Data

The code chunk below imports the data into R and saves it into a tibble data frame:

```{r}
covid19 <- read_csv("data/COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)

```

# 2.0 `FunnelPlotR` Methods

## 2.1 `FunnelPlotR` Methods: The Basic Plot
