---
title: "Hands-on Exercise 01: A Layered Grammar of Graphics: ggplot2 Methods"
author: "Sheryl Ann Tan Yi-Shi"
date-modified: "last-modified"
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

## Loading the Required Libraries

```{r}
pacman::p_load(tidyverse)
```

## Importing Data

The code chunk below imports exam_data.csv into R environment by using the read_csv() function of the readr package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

There data set contains seven attributes:

-   **Categorical Attributes**: ID, CLASS, GENDER and RACE. These columns are denoted as *chr* ("character type").

-   **Continuous Attributes**: ENGLISH, MATHS, and SCIENCE. These columns are stored as *dbl* ("double type"), i.e. numeric values.

## R Graphics vs. ggplot

The tabset below compares how R Graphics and ggplot plot a simple histogram:

# Histogram Comparison

::: panel-tabset
## R Graphics

```{r}
hist(exam_data$MATHS)
```

## ggplot2

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```
:::
