---
title: "Hands-on Exercise 5.5: Treemap Visualisation with R"
author: "Sheryl Ann Tan Yi-Shi"
date: "2025-02-08"
date-modified: "last-modified"
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1.0 Getting Started

## 1.1 Installing and Launching R Packages

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

## 1.2 Importing the Data

The `read_csv()` is used to import *realis2018.csv* into R:

```{r}
realis2018 <- read_csv("data/realis2018.csv")
```

## 1.3 Data Wrangling and Manipulation

As *realis2018* is in transaction record form, which is highly disaggregated and not appropriate to be used to plot a treemap.

### 1.3.1 Grouped Summaries without the Pipe

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

### 1.3.2 Grouped Summaries with the *Pipe*

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

# 2.0 Designing Treemap with `treemap` Package

## 2.1 Designing a Static Treemap

The `treemap()` function of the **Treemap** package is used to plot a treemap showing the distribution of median unit prices and total unit sold of resole condominium by geographic hierarchy.

The following code selects records belonging to resale condominium:

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

## 2.2 Using Basic Arguments

The code chunk below creates a treemap using the following argument: *index, vSize and vColor*.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

**Note:**

-   **Index:** The index vector must consist of at least two column names or else no hierarchy treemap will be plotted.

-   **vSize**: The column must not contain negative values.

## 2.3 Working with *vColor* and *type* Arguments

The *type* argument is defined as value in the code chunk below:

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type = "value",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 2.4 The "Value" Type Treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 2.5 The "Manual" Type Treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

The following code chunk uses a single colour palette instead of diverging colour palette. Single colour palette should be used in cases where the values are **all** positive or negative:

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 2.6 Working with the *algorithm* Argument

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "squarified",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## 2.7 Using *sortID*

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

# 3.0 Designing Treemap Using `treemapify` Package

## 3.1 Designing a Basic Treemap

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`),
       layout = "scol",
       start = "bottomleft") + 
  geom_treemap() +
  scale_fill_gradient(low = "light blue", high = "blue")
```

## 3.2 Defining Hierarchy

Group by Planning Region

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```

**Group by Planning R**

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap()
```

**Adding Boundary Line**

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```

# 4.0 Designing Interactive Treemap Using `d3treeR`

## 4.1 Installing the `d3treeR` Package

```{r}
library(d3treeR)
```

## 4.2 Designing an Interactive Treemap

The code uses *`treemap()`* to build a treemap by using selected variables in condominium data.frame.

```{r}
tm <- treemap(realis2018_summarised,
        index=c("Planning Region", "Planning Area"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        title="Private Residential Property Sold, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

`d3tree()` is then used to build an interactive treemap:

```{r}
d3tree(tm,rootname = "Singapore" )
```
