---
title: "Hands-on Exercise 5.4: Visual Multivariate Analysis with Parallel Coordinates Plot"
author: "Sheryl Ann Tan Yi-Shi"
date: "2025-02-08"
date-modified: "last-modified"
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# 1.0 Getting Started

Parallel coordinates plot compares multiple variables together and seeing the relationships between. This visualisation technique is more often found in academic and scientific studies rather than business / consumer data visualisations.

Parallel coordinates is used to bring meaningful multivariate patterns and comparisons to light.

## 1.1 Installing and Launching R Packages

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

## 1.2 Importing Data

The `read_csv()` function is used to import *WHData-2018.csv* into R:

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

# 2.0 Plotting Static Parallel Coordinates

## 2.1 Plotting a Simple Parallel Coordinates Plot

The code chunk below uses `ggparcoord()` to plot a basic static parallel coordinates plot:

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

## 2.2 Plotting a Parallel Coordinates with Boxplot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

## 2.3 Parallel Coordinates with Facet

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

## 2.4 Rotating X-Axis Text Label

The `theme()` function in `ggplot2` is used to rotate the text labels:

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

## 2.5 Adjusting the Rotated X-Axis Text Label

`hjust` is used to prevent label overlap with the plot:

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

# 3.0 Plotting Interactive Coordinates Plot: `parallelPlot` Methods

## 3.1 The Basic Plot

The code chunk uses `parallelPlot()` to plot an interactive parallel coordinates plot:

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

## 3.2 Rotate Axis Label

The axis label is rotated using the `rotateTitle` argument:

```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

## 3.3 Changing the Colour Scheme

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

The code provided above did not change the colour scheme despite including the argument `continuousCS`. We would need to included the `refColumnDim` argument ensure that the continuousCS effect shows:

(Ref: <https://cran.r-project.org/web/packages/parallelPlot/vignettes/introduction-to-parallelplot.html>)

```{r}
parallelPlot(wh,
             refColumnDim = "Happiness score",
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

## 3.4 Parallel Coordinates Plot with Histogram

The `histoVisibility` argument is used to plot histogram along the axis of each variable:

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```

# 4.0 References

-   [*ggparcoord()*](http://ggobi.github.io/ggally/reference/ggparcoord.html) of [**GGally**](http://ggobi.github.io/ggally/index.html) package

-   [**parcoords** user guide](https://cran.r-project.org/web/packages/parcoords/parcoords.pdf)

-   [**parallelPlot**](https://cran.r-project.org/web/packages/parallelPlot/parallelPlot.pdf)
